Stability of variational quantum dynamics with Gaussian approximants

### 2013-06-28 paper.tex
figured out SVE of coherent state expansion.

@@

>>>>

The variational methods are based on the phase space mapping $G$ from square-integrable functions to Fock space, where
$$Gf=∫f(α*)|α〉\,d²α.$$
This is different from Glauber's use: my $f$ is his $fe^{-½|α|²}$.  (Check the half.)

Hypothesis: The singular values of $G$ are 0 and $\sqrt π$, and the space belonging to $\sqrt π$ comprises the functions $f$ for which $fe^{½|α|²}$ is analytic.

There are two ways one could set about proving this.  Firstly, one could show that any function orthogonal to the Gaussian analytics annihilates $G$.  Otherwise, one could show that the norms of all Gaussian analytics are scaled by the same factor going to F
ock space, and that they are a pre-image of Fock space.  The second approach sounds simpler.

This won't be a rigourous argument: it could turn out that there's another class of expansions with half the norm, and the entire functions have components belonging to the 0 singular value.  My numerical experiment suggests otherwise; let's ignore that possibility for now.

Let $g$ be a normalised entire function, in Glauber's sense.  Our expansion function is 
$$f(α) = {1\over π}g(α)e^{-½|α|²}.$$
If $g$ has the Taylor expansion
$$g(z)=∑_n c_n{zⁿ\over(n!)^½},$$
then 
$$Gf = ∑_n c_n|n〉.$$
The functions $zⁿ$ for $n≥0$ span the entire functions (or their span is a dense subset, or something like that), so let's find the norm of 
$$f(α)={1\over π}e^{-½|α|²}{αⁿ\over(n!)^½}.$$
We have 
$$|f(α)|²={1\over π²n!}e^{-|α|²}|α|^{2n},$$
so the squared norm is
$$\|f\|²={1\over π²n!}∫₀^∞2πr^{2n+1}e^{-r²}\,dr={2\over πn!}∫₀^∞r^{2n+1}e^{-r²}\,dr.$$
According to Wolfram, this is 
$$\|f\|²={1\over π},$$
whence
$$\|f\|=π^{-½}.$$

Note that the gaussian entire functions are a vector subspace of the square integrable functions.

Assuming that there is no way to expand a state with a smaller norm than a Gaussian entire function, this does it.  Or at least I think so.  I should look up exactly what the uniqueness properties are for singular value expansions in infinite dimensional spaces.

@@

The things I've noticed so far are:
\item{1.} When a superposition is iteratively fitted to the exactly propagated state, adding a term to the cost proportional to the Jacobian $|Dψ〉$ keeps the fit stable for an entire cycle.  It isn't accurate, though.

\item{2.} Measuring state differences in a truncated fock space makes the fit stable and accurate.  Of course, the truncated state is different from the original one, so the original state is not fit precisely.  When this was done, the least squares problem involved 6 fock coefficients, and the amplitude vector has 10 elements.  So almost all the truncated problems were underdetermined: some singular values of the Jacobian are exactly zero.  Matlab claims that backslash solves underdetermined problems by giving a solution with as many zero elements as possible.  The documentation doesn't say how it chooses which elements to make zero.  That's different from the other regularisation methods we've tried.  Of course, the problems are being solved inside Matlab's optimisation routine, and I don't know what it's doing.

Reducing the dimension of Fock space so that the problem is underdetermined means that, when the problem is solved by SVD, we only take components along $N+1$ directions in Fock space, and map them to $N+1$ directions in parameter space.  The remaining $2R-N-1$ components of the solution are zero.  As we add extra dimensions to Fock space, we are taking components in more Fock space directions, which include more oscillating components of the wavefunction.  This is like sampling a function on a finer grid.  With these directions, we can form some of the smaller singular vectors of the full problem.

Truncating Fock space has a similar effect to truncating the singular values, assuming the singular directions belonging to small values have large particle numbers.  (And they do: when I plotted their overlaps with coherent states, they formed rings without support at the origin.)  And indeed, the trajectories for the average amplitude change in similar ways due to truncation: they pick up an extra loop in the complex plane.

As the Fock space expands with $N»2R$, there are still only $2R$ singular directions.  Presumably these converge to limits as $N$ increases.  There might be a norm on the matrices $〈N|Dψ〉$, where the truncated elements replaced by zero, that measures this.

\item{3.} The difference between the truncated state and the original one fails to satisfy the Picard condition when with respect to $|Dψ〉$.  What about $H|ψ〉$?

\item{4.} When the least squares problem is solved with a different distance function, which sets distances along the large $N$ directions to zero, the fit fails on the first step.  Note that this isn't really a distance.  Also, I did it by solving $P|h〉=P|Dψ〉dz$; when $P$ truncates the Fock space, the operator $P|Dψ〉$ is singular.  Doing a weighted least squares fit, with the many particle directions having low weights, would be near-singular.
